A Comparison of Dimensionality Reduction Techniques for Robotic Motion Planning

### 2016-05-13 paper/paper.tex
updates to paper

Principle Component Analysis, or PCA, transforms a set of variables into their
``principle components", or linearly uncorrelated variables. In other words,
the axes of the dataset in question are reoriented so the Rayleigh quotient,
which is related to the covariance of the dataset, along the first axis is
maximized. After the first axis has been set, the second axis is then set to
maximize the Rayleigh quotient of the dataset along that axis, and so on, so
that the first axis has the highest Rayleigh quotient, the second axis has the
second highest, and so on. Dimensionality reduction is then performed by
chopping off the axes with the lowest Rayleigh quotients (which corresponded to
axes with the lowest variance).

$$
\mathbf{w}_{(1)} = {\operatorname{\arg\,max}}\, \left\{ \frac{\mathbf{w}^T\mathbf{X}^T \mathbf{X w}}{\mathbf{w}^T \mathbf{w}} \right\}
$$

@@

In Randomized PCA, computations are sped up by only approximating the singular vectors of the components.

@@

Feature Agglomeration groups together similar features to reduce the number of
features. It does this by using agglomerative clustering, in which every data
point starts as a cluster, and clusters are merged together.

@@

Truncated SVD works best with sparse matrices.

>>>>

Principle Component Analysis, or PCA, transforms a set of variables into their
``principal components", or linearly uncorrelated variables \cite{shlens2014pca}. Given an $m \times n$ matrix $\mathbf{X}$, you can find its principal components by finding an orthonormal matrix $\mathbf{P}$ in $ \mathbf{Y} = \mathbf{PX} $ such that $\mathbf{C_Y} = \mathbf{YY}^T/n$. The rows of $\mathbf{P}$ are the principal components of $\mathbf{X}$. In other words,
the axes of the dataset in question are reoriented so that the covariance of the data along the first axis is minimized. After the first axis has been set, the second axis is then set to
minimize the covariance of the dataset along that axis, and so on, so
that the first axis has the lowest covariance, the second axis has the
second lowest, and so on. Dimensionality reduction is then performed by
chopping off the axes with the high covariance.

@@

In Randomized PCA, computations are sped up by approximating the largest singular vectors of the components.

@@

Feature Agglomeration is a form of hierarchical clustering in which clusters are merged or split successively \cite{featureagglomeration}. Feature Agglomeration groups together similar features to reduce the number of
features. It does this by using agglomerative clustering, in which every data
point starts as a cluster, and clusters are merged together.

@@

In singular value decomposition, a matrix $\mathbf{M}$ is factorized into the form $\mathbf{U} \Sigma \mathbf{V^*}$. $\Sigma$ is a diagonal matrix, and the entries of $\Sigma$ are the singular values of $\mathbf{M}$. In Truncated SVD, $\mathbf{M}$ is approximated by matrices $\mathbf{U_k} \Sigma_k \mathbf{V^*_k}$, where $ k < r $, $r$ being the number of non-zero singular values of the matrix \cite{maciejewski1989svd}. In effect you are approximating $\mathbf{M}$ by cutting off low-value singular values. In this sense Truncated SVD is similar to PCA; however, instead of cutting off the lowest eigenvalues of the coviarance matrix, you are cutting off the singular values of the matrix.
